---
title: "Figure3_CMP"
output: html_document
date: "2024-05-28"
---

# library
```{r}
library("sva")
library("ggplot2")
library("gridExtra")
library("edgeR")
library("UpSetR")
library("tidyverse")
library("BatchQC")
library("RColorBrewer")
library("Glimma")
library("variancePartition")
library("BiocParallel")
library("gplots")
library("NMF")
library("GO.db")
library("org.Hs.eg.db")
library(preprocessCore)
library(genefilter)
library(matrixStats)
library(rafalib)
library(PCAtools)
library(Biobase)
library(GEOquery)
library(bigutilsr)
library(spatstat)
library(matrixStats)
library(RCurl)
library(GSA)
library(RCy3)
library(fgsea)
library(data.table)
# library(plotly) # has conflict with org.Hs.eg.db select
# library(rrcov) # has conflict with biplot detach(rrcov)
library(DT) 
library(grid)
library(ComplexHeatmap)
library(cluster)

```

```{r}
load("~/JH/HF/Spatial_original_basedata.RData")
```

```{r}
# ##################################################################
# 'TMA.block'
# 'ROI.size'
# 'Ventricle'
# 'Segment.id'
# 'Segment.type' -> "celltype"
# 'PID'
# 'Operation'
# 'Fixation.interval'
# 'Gender'
# 'Age'
# 'Institute'
# 'Clinical_phenotype_path'
# 'Clinical_endstage'
# 'After_LVAD'
# 'P_Normal'
# 'P_Degeneration''P_Hypertrophy''P_Disarray''P_Fibrosis''P_peri_infarction''P_etc'
# 'Echo_date''Echo_LVEF''Echo_LVPWd''Echo_IVSd''Echo_chamber_size''Echo_RWMA''Echo_RV_dysfunction''Echo_AR''Echo_MR''Echo_TR'
# 'ECG_date''ECG_voltage''ECG_QRS''VTandVF''AF''BNP''Probnp''CardiacMR_DE''Gene'
# 'ECG_voltage_n''ECG_QRS_n'
# 'Core.c.fullROI'
# 'PID.Op''PID.Op.bothblock''ROI.bothsize'
# 'Clinical_phenotype_LV''Clinical_phenotype_tri''Clinical_phenotype_bi'


# Number of core to use for DREAM analysis
 NoCore = 8 # 8

# Normalization method 
 nlm = "upperquartile" # "upperquartile"

# # Condition of interest (among column names of met)
  # COI <- "Clinical_phenotype_bi"   # COI <- "celltype" 
    COI <- "celltype" 
# USING ONLY SPECIFIC SET of CORES - For example cores with full ROI 
  # FullROI_selection <- "Y"
  FullROI_selection <- "N"

# CELL TYPE SELECTION?
  celltype_selection <- "N"  # or "N"
  # celltype_selection <- "Y"  # or "N"

  # Which celltype? 
  cell_type<- "Cardiomyocytes" # or ...

# ROISIZE SELECTION?
 ROISIZE_selection <- "Y" # or "N"
 ROIsize <- "large" # or ...

# VENTRICLE SELECTION?
    ventricle_selection <- "N"
    # ventricle_selection <- "Y" # or "N"
    ventricle_type <- "LV"

# Clinical Disease selection with control samples

    # disease_selection <- "Y"
    disease_selection <- "N"

    disease_selection_var <- "Clinical_phenotype_LV"
    disease_type_vector <- c("Control", "ES_HCMP", "NES_HCMP")  ## Clinical_phenotype_LV - 'ES_HCMP''ICMP''Control''DCMP''NES_HCMP'
    disease_type <- if(disease_selection == "Y") {disease_type = disease_type_vector
                                                  }else{disease_type= "Non"
                                                       }

# OUTLIER SELECTION?
 # if no 
  grid_set = "Y" 
  # grid_set = "N"

 # choose which grid set to use from the PCAgrid()
if (grid_set == "N") {
    outlier_set = "N"
    } else {
    outlier_set =  "pcaplot_outlier" #"pca_grid_10"# "pca_grid_10" # or pca_grid_10, pca_grid_3, etc
}

RE1 = "PID"
RE2 = "Institute"
RE3 = "Fixation.interval"
# RE4 = "P_Normal"
# RE4 = "Clinical_phenotype_LV"
# RE4 = "..." 
#...

if (exists("RE4")==TRUE) {
    (form <- formula(paste("~", "condition + (1|", RE1, ") + (1|", RE2, ") + (1|", RE3, ") + (1|", RE4, ")", collapse="")))
    (form_cont <- formula(paste("~ 0 +", "condition + (1|", RE1, ") + (1|", RE2, ") + (1|", RE3, ") + (1|", RE4, ")", collapse="")))
# RE names
    (RE_names <- paste(RE1, RE2, RE3, RE4, sep="_"))
    }else{
    if (exists("RE3")==TRUE) {
        (form <- formula(paste("~", "condition + (1|", RE1, ") + (1|", RE2, ") + (1|", RE3, ")", collapse="")))
        (form_cont <- formula(paste("~ 0 +", "condition + (1|", RE1, ") + (1|", RE2, ") + (1|", RE3, ")", collapse="")))

# RE names
        (RE_names <- paste(RE1, RE2, RE3, sep="_"))
        }else{
        if (exists("RE2")==TRUE) {
            (form <- formula(paste("~", "condition + (1|", RE1, ") + (1|", RE2, ")", collapse="")))
            (form_cont <- formula(paste("~ 0 +", "condition + (1|", RE1, ") + (1|", RE2, ")", collapse="")))

# RE names
            (RE_names <- paste(RE1, RE2, sep="_"))
            }else{
            if (exists("RE1")==TRUE) {
                (form <- formula(paste("~", "condition + (1|", RE1, ")", collapse="")))
                (form_cont <- formula(paste("~ 0 +", "condition + (1|", RE1, ")", collapse="")))

# RE names
                (RE_names <- paste(RE1, sep="_"))
                }else{
                (form <- formula(paste("~", "condition", collapse="")))
                (form_cont <- formula(paste("~ 0 +", "condition", collapse="")))

# RE names
                (RE_names <- paste(No, sep="_"))
                }
        }
    }
}
                
# Decide Contrast for DREAM

#makecontrastasroutine <- "celltype_comwith_othercells" 
# makecontrastasroutine <- "Forclinicaldisease" # or "Y" to make it according to the number of condition
# makecontrastasroutine <- "Y" #to make it according to the number of condition

# Small sample number analysis 
 KRanaly <- "Y"
 # KRanaly <- "N"
# Order of batch correction
# Cell-type selection -> Batch correction

# Outlier 
# remove Sejong
# Normalization method = Upper quartile 
# ##################################################################
```

```{r}
(Analysis_name = paste("SejongIn_", COI, "_FullROI_", FullROI_selection, 
                      "_cellselection_", celltype_selection,
                                "_", cell_type, 
                                "_ROI_", ROIsize,
                                "_Vent_", ventricle_type,
                                "_DsSelect_", paste0(disease_type, "_", collapse="_"), disease_selection, 
                                "_", "OL_", outlier_set,
                                "_RE_", RE_names,
                                "_KR_", KRanaly,
                      Sys.Date(), sep=""))
```

# INPUT DATA

##  Input data - new 

```{r}
# Decide Input file
Input_expr_data = "Probe_QC_1percfilter.csv"
Input_pheno_data = "Spatial_Annotation.csv"

# Read Input files
exp_qc <- read.csv(Input_expr_data, row.names = 1, header = TRUE, stringsAsFactors=FALSE)
met_tmp<- read.csv(Input_pheno_data, header=TRUE, stringsAsFactors=FALSE)
```

```{r}
# Refine met file to match expression file
met_tmp1 <- met_tmp %>% mutate(Name=paste(Scan.name, formatC(ROI..label., width=3, flag="0"), Segment..Name..Label., sep="..."))
rownames(met_tmp1)<- gsub(' ', '.', met_tmp1$Name)

met<-met_tmp1[match(colnames(exp_qc), rownames(met_tmp1)),]
identical(colnames(exp_qc), rownames(met))

met <- met %>% mutate(TMA.block = factor(TMA.block))
```

```{r}
# define celltype
celltype <- as.factor(met$Segment..Name..Label.)
celltype <- gsub("Full ROI", "No_Seg",celltype) 
celltype <- gsub("Trp", "CMC", celltype)
celltype <- gsub("vessel", "EC", celltype)
celltype <- gsub("fb", "FB", celltype)
```

```{r}
met <- cbind(met, celltype)
```

## Remove negative probe

```{r}
negprobe=which(rownames(exp_qc)%in%"NegProbe-WTX")

class(negprobe)

exp_rmnp <- exp_qc[-negprobe,]
dim(exp_rmnp)
dim(exp_qc)
```

## GENE ANNOTATION (GENEID)

```{r}
if ("plotly" %in% tolower((.packages())) == TRUE) {
    detach("package:plotly", unload=TRUE)
    }
```

```{r}
# select makes error with clusterProfiler
if ("clusterProfiler" %in% (.packages()) == TRUE) {
detach("package:clusterProfiler", unload=TRUE)
    }
```

```{r}
symbols <- rownames(exp_rmnp)
```

```{r}
fdata <- select(org.Hs.eg.db, symbols, c("ENTREZID", "GENENAME"), "SYMBOL")
```

```{r}
identical(fdata$SYMBOL,symbols)
```

```{r}
## There are duplicates and missing
```

```{r}
## first remove duplicates
```

```{r}
## MENO1 take 51072 according to nanostring excel information
```

```{r}
which(fdata$SYMBOL=="MEMO1")
```

```{r}
fdata_nodup <- fdata[-17,]
```

```{r}
dim(fdata_nodup)
```

```{r}
which(fdata_nodup$SYMBOL=="MEMO1")
```

```{r}
## Replace Missing ENTREZIDs with IDs in Geomx file
```

```{r}
symbols_noID<-fdata_nodup$SYMBOL[which(is.na(fdata_nodup$ENTREZID))]
length(symbols_noID)
```

```{r}
gene_list = "gene_list.csv"
geomx_gene <-read.csv(gene_list, header=TRUE, stringsAsFactors=FALSE)
```

```{r}
geomx_glist<-geomx_gene[,c("TargetName","GeneID")]
```

```{r}
missing_geomx <- geomx_glist$TargetName[which(geomx_glist$TargetName%in%symbols_noID)]
length(which(geomx_glist$TargetName%in%symbols_noID))
```

```{r}
missing_glist<-geomx_glist[which(geomx_glist$TargetName%in%symbols_noID),]
```

```{r}
reorder_missing_glist<-missing_glist[match(symbols_noID,missing_geomx),]
identical(reorder_missing_glist$TargetName, symbols_noID)

# fdata_nodup$ENTREZID[which(is.na(fdata_nodup$ENTREZID))]<-missing_geomx_tb$GeneID
```

```{r}
fdata_nodup$ENTREZID[which(is.na(fdata_nodup$ENTREZID))] <- reorder_missing_glist$GeneID
```

### check duplicated IDs

```{r}
test_dupl<-fdata_nodup[order(fdata_nodup$ENTREZID),]
head(test_dupl)
```

```{r}
which(duplicated(test_dupl$ENTREZID)==TRUE)
```

```{r}
duplicated_fdata<-test_dupl[duplicated(test_dupl$ENTREZID)|duplicated(test_dupl$ENTREZID, fromLast=TRUE),]
```

```{r}
duplicated_fdata
```

```{r}
## confirmed that there is no duplicates
```

### If wants to change rownames to ENTREZID

```{r}
# rownames(e_abc)<-fdata_nodup$ENTREZID
```

## Three data sets

```{r}
Input_for_bc = exp_rmnp
```

```{r}
e_bbc <- Input_for_bc # before batch correction 

f <- fdata_nodup
p <- met
```

```{r}
unique(p$Segment.id)
```

```{r}
identical(f$SYMBOL,symbols)
```

```{r}
d_p <- DGEList(e_bbc)
d_pq3 <- calcNormFactors(d_p, method="upperquartile")
d_pq3cpm <- cpm(d_pq3, log=FALSE)
```

# GROUP SELECTION & BATCH CORRECTION

## Select Full-ROI set

```{r}
 X0="Core.c.fullROI"
 Y0="Yes"
 W0="p"
 Z0="e_bbc" 
if (FullROI_selection == "Y") {

select_group(X0,Y0,W0,Z0)

output_p0 = get(paste(W0, "_", Y0, sep=""))
output_e0 = get(paste(Z0, "_", Y0, sep=""))
    } else {
    output_p0 = get(W0)
    output_e0 = get(Z0)
    }
```

## Select Cell-type 

```{r}
 X1="Segment.id"
 Y1= cell_type
 W1="output_p0"
 Z1="output_e0"

if (celltype_selection == "Y") {

select_group(X1,Y1,W1,Z1)

output_p1 = get(paste(W1, "_", Y1, sep=""))
output_e1 = get(paste(Z1, "_", Y1, sep=""))
    } else {
    output_p1 = get(W1)
    output_e1 = get(Z1)
    }
```

## Batch Correction (TMA.block)

### BatchQC before batchcorrection

```{r}
output_e1_prenormal <- newSeqExpressionSet(as.matrix(output_e1), phenoData = output_p1)
output_e1_q3 <- betweenLaneNormalization(output_e1_prenormal, which ="upper")
```

```{r}
## BatchQC before correcting batch effect
batch = output_p1$TMA.block
condition = output_p1$Segment.type
batchQC(dat=normCounts(output_e1_q3), batch=batch, condition=condition,
        report_file="Supplemental_Figure1_beforebatchcorrection_20230621.html", report_dir=".",
        report_option_binary="111111111",
        view_report=TRUE, interactive=TRUE, batchqc_output=TRUE)
```

### Batch correction

```{r}
# select COI
condition = output_p1[[COI]]
```

```{r}
# Input data
Input_for_bc_2 = output_e1
#perform the batch correction
batch = output_p1$TMA.block
##** need to correct condition according to the phenotype interesting
# condition = p_Cardiomyocytes$P_Normal

#groups = sapply(as.character(condition), switch, "Full ROI" = 1, "Trp" = 2, "vessel" = 3, "fb" = 4, USE.NAMES = F)
batches = sapply(as.character(batch), switch, "1" = 1, "2" = 2, USE.NAMES = F)
output_bc = ComBat_seq(counts = as.matrix(Input_for_bc_2), batch = batches, group = condition)
```

```{r}
### BatchQC after batchcorrection
```

```{r}
output_bc_prenormal <- newSeqExpressionSet(as.matrix(output_bc), phenoData = output_p1)
output_bc_q3 <- betweenLaneNormalization(output_bc_prenormal, which ="upper")
```

```{r}
## BatchQC before correcting batch effect
batch = output_p1$TMA.block
condition = output_p1$Segment.type
batchQC(dat=normCounts(output_bc_q3), batch=batch, condition=condition,
        report_file="Supplemental_Figure1_afterbatchcorrection_20230621.html", report_dir=".",
        report_option_binary="111111111",
        view_report=TRUE, interactive=TRUE, batchqc_output=TRUE)
```

## Select Large and LV only ROI.size=='large'

```{r}
    X2="ROI.size"
    Y2= ROIsize
    W2="output_p1"
    Z2="output_bc"
## select Large and LV only ROI.size=='large'
if (ROISIZE_selection == "Y") {
    
    select_group(X2,Y2,W2,Z2)
    
    output_p2 = get(paste(W2, "_", Y2, sep=""))
    output_e2 = get(paste(Z2, "_", Y2, sep=""))
} else {
    output_p2 = get(W2)
    output_e2 = get(Z2)
}
```

## Select Large and LV only Ventricle=='LV'

```{r}
    X3="Ventricle"
    Y3=ventricle_type
    W3="output_p2"
    Z3="output_e2"
# input = output of above cell

if (ventricle_selection == "Y") {
    select_group(X3,Y3,W3,Z3)
    
    output_p5 = get(paste(W3, "_", Y3, sep=""))
    output_e5 = get(paste(Z3, "_", Y3, sep=""))
    } else {
    output_p5 = get(W3)
    output_e5 = get(Z3)
    }
```

## Select Specific Disease and Control samples

```{r}
X5=disease_selection_var
Y5="disease_type"
W5="output_p5"
Z5="output_e5"

if (disease_selection == "Y") {
    select_group_m(X5,Y5,W5,Z5)
    
    output_p3 = get(paste(W5, "_", Y5, sep=""))
    output_e3 = get(paste(Z5, "_", Y5, sep=""))
    } else {
    output_p3 = get(W5)
    output_e3 = get(Z5)
    }
```

# Intermediate DATA INPUT

```{r}
d_1 <- DGEList(output_e3)
```

### Add normalization factors

```{r}
d_1q3 <- calcNormFactors(d_1, method="upperquartile")
# d_1tmm <- calcNormFactors(d_1)
```

```{r}
dim(d_1)
```

```{r}
head(d_1)
```

# PCA ANALYSIS

```{r}
if ("rrcov" %in% tolower((.packages())) == TRUE) {
    detach("package:rrcov", unload=TRUE)
    }
```

```{r}
d_1q3cpm<-cpm(d_1q3, log=FALSE)
```

```{r}
e.list <- list(d_1q3cpm)
```

```{r}
pca.list <- lapply(e.list, function(x) pca(x, metadata=output_p3, removeVar = 0.1, scale = TRUE))
```


# OUTLIER

## Outlier removal according to PCA plot

```{r}
pca2_tmp<-pca.list[[1]]$rotated[2]
pcaplot_outlier <-rownames(pca2_tmp)[pca2_tmp$PC2>(50)]

pcaplot_outlier
```

```{r}
dim(output_e3)
```

```{r}
Y4="output_p3"
Z4="output_e3"

if (grid_set == "N") {
    output_p4 = get(Y4)
    output_e4 = get(Z4)
    } else {
    if (grepl("pcaplot", outlier_set)==TRUE) {
        X4=outlier_set
        } else {
        grid_set = get(outlier_set)
        sp_gridflag <-names(grid_set$flag[grid_set$flag==FALSE])
        X4="sp_gridflag"
        }
    remove_sample(X4,Y4,Z4)
    output_p4 = get(paste(Y4, "_rm_", X4, sep=""))
    output_e4 = get(paste(Z4, "_rm_", X4, sep=""))
    }
```

# FINAL DATA INPUT

```{r}
d0 <- DGEList(output_e4)
m <- output_p4
e <- output_e4
f <- fdata_nodup
d0$genes <- f
```

### Add features of samples in "samples"

```{r}
d0$samples$celltype <- m$celltype

PID <- as.factor(m$PID)
d0$samples$PID <- PID

condition <- as.factor(m[[COI]])
d0$samples$condition <- condition

### remove after confirmation

# Ds <- as.factor(m$Clinical_phenotype_2nd)
# d0$samples$Ds <- Ds

# levels(condition) <- c("Non_Endstage", "End_stage")
m <- cbind(m, condition)
```

### Add normalization factors

```{r}
d1 <- calcNormFactors(d0, method="upperquartile")
```

# Heatmaps and PCA plot after all

## Heatmaps with lcpm

```{r}
logcounts <- cpm(d1, log=TRUE)
```

# Figure 3b
```{r}
nature_cvr <- load("~/HF/HF_cvr/GSE183852_DCM_Integrated.Robj")

Idents(nature_cvr) <- "tech"
HF_cvr_sn <- subset(nature_cvr, idents = "SN")

Idents(HF_cvr_sn) <- "Names"

HF_cvr_sn2 <- subset(HF_cvr_sn, idents = c("Cardiomyocytes", "Endothelium", "Fibroblasts", "Pericytes", "Smooth_Muscle", "Myeloid", "NK/T-Cells", "B-Cells", "Neurons", "Lymphatic", "Mast", "Adipocytes"))

my_cvr_order <- c("Cardiomyocytes", "Endothelium", "Fibroblasts", "Pericytes", "Smooth_Muscle", "Myeloid", "NK/T-Cells", "B-Cells", "Neurons", "Lymphatic", "Mast", "Adipocytes")

Idents(HF_cvr_sn2) <- "Names"
DefaultAssay(HF_cvr_sn2) <- "RNA"
cmc <- subset(HF_cvr_sn2, idents = c("Cardiomyocytes"))
endo <- subset(HF_cvr_sn2, idents = c("Endothelium"))
fb <- subset(HF_cvr_sn2, idents = c("Fibroblasts"))

Idents(cmc) <- "condition"
Idents(endo) <- "condition"
Idents(fb) <- "condition"

cmc <- NormalizeData(cmc)
endo <- NormalizeData(endo)
fb <- NormalizeData(fb)

cmc_marker <- RunPrestoAll(cmc, logfc.threshold = 0.25)
endo_marker <- RunPrestoAll(endo, logfc.threshold = 0.25)
fb_marker <- RunPrestoAll(fb, logfc.threshold = 0.25)

endo_dcmp_marker <- endo_marker[endo_marker$cluster == "DCM",]
endo_dcmp_marker <- endo_dcmp_marker[endo_dcmp_marker$p_val_adj < 0.05,]
endo_dcmp_marker_gene <- endo_dcmp_marker$gene

cmc_dcmp_deg <- cmc_marker[cmc_marker$cluster == "DCM",]
cmc_dcmp_deg <- cmc_dcmp_deg[cmc_dcmp_deg$p_val_adj < 0.05,]
cmc_dcmp_deg_gene <- cmc_dcmp_deg$gene

fib_dcmp_deg <- fb_marker[fb_marker$cluster == "DCM",]
fib_dcmp_deg <- fib_dcmp_deg[fib_dcmp_deg$p_val_adj < 0.05,]
fib_dcmp_deg_gene <- fib_dcmp_deg$gene
```

```{r}
cmc_roi_dcmp_deg <- read.csv("~/JH/HF/RData/results/SejongIn_Clinical_phenotype_LV_FullROI_N_cellselection_Y_Cardiomyocytes_ROI_large_Vent_LV_DsSelect_DCMP__Control_Y_OL_pcaplot_outlier_RE_PID_Institute_Fixation.interval_KR_Y2023-06-22conditionDCMP-conditionControl.csv")
cmc_roi_dcmp_deg <- cmc_roi_dcmp_deg[cmc_roi_dcmp_deg$adj.P.Val < 0.05,]
cmc_roi_dcmp_deg_gene <- cmc_roi_dcmp_deg$X
endo_roi_dcmp_deg <- read.csv("~/JH/HF/RData/results/SejongIn_Clinical_phenotype_LV_FullROI_N_cellselection_Y_Endothelial_cells_ROI_large_Vent_LV_DsSelect_DCMP__Control_Y_OL_pcaplot_outlier_RE_PID_Institute_Fixation.interval_KR_Y2023-06-23conditionDCMP-conditionControl.csv")
endo_roi_dcmp_deg <- endo_roi_dcmp_deg[endo_roi_dcmp_deg$adj.P.Val < 0.05,]
endo_roi_dcmp_deg_gene <- endo_roi_dcmp_deg$X

## Validation celltype ----
geomx_gene <- endo_roi_dcmp_deg$X
nature_gene <- rownames(HF_cvr_sn2@assays$RNA@counts)

#Intersect all variable genes
DefaultAssay(HF_cvr_sn2) <- "RNA"
common.genes <- intersect(nature_gene, geomx_gene)
```


```{r}
dcmp_gene <- c(cmc_roi_dcmp_deg_gene, endo_roi_dcmp_deg_gene, cmc_dcmp_deg_gene, endo_dcmp_marker_gene, fib_dcmp_deg_gene)
dcmp_gene <- unique(dcmp_gene)
dcmp_gene <- intersect(dcmp_gene, nature_gene)
geomx_gene <- rownames(avr.geo)
dcmp_gene <- intersect(dcmp_gene, geomx_gene)

dcmp <- subset(HF_cvr_sn2, idents = c("DCM"))

dcmp$Names <- factor(dcmp$Names,
                         levels = c("Cardiomyocytes", "Endothelium", "Fibroblasts", "Pericytes", "Smooth_Muscle", "Myeloid", "NK/T-Cells", "B-Cells", "Neurons", "Lymphatic", "Mast", "Adipocytes"))

avr.geo <- read.csv("~/JH/HF/logcpmnorm.csv")
rownames(avr.geo) <- avr.geo$X
avr.geo <- avr.geo[,2:158]
geo.roi <- colnames(avr.geo)
metadata <- read.csv("~/JH/HF/RData/Spatial_Annotation.csv")
metadata <- metadata[metadata$X %in% geo.roi,]
metadata_cmc <- metadata[metadata$Segment..Name..Label. == "Trp",]
metadata_cmc_dcmp <- metadata_cmc[metadata_cmc$Clinical_phenotype_LV == "DCMP",]
metadata_cmc_dcmp_roi <- metadata_cmc_dcmp$X
metadata_endo <- metadata[metadata$Segment..Name..Label. == "vessel",]
metadata_endo_dcmp <- metadata_endo[metadata_endo$Clinical_phenotype_LV == "DCMP",]
metadata_endo_dcmp_roi <- metadata_endo_dcmp$X
metadata_fb <- metadata[metadata$Segment..Name..Label. == "fb",]
metadata_fb_dcmp <- metadata_fb[metadata_fb$Clinical_phenotype_LV == "DCMP",]
metadata_fb_dcmp_roi <- metadata_fb_dcmp$X

avr.geo.dcmp.cmc <- avr.geo[dcmp_gene, metadata_cmc_dcmp_roi]
avr.geo.dcmp.cmc$cmc_avg <- rowMeans(avr.geo.dcmp.cmc)
avr.geo.dcmp.endo <- avr.geo[dcmp_gene, metadata_endo_dcmp_roi]
avr.geo.dcmp.endo$endo_avg <- rowMeans(avr.geo.dcmp.endo)
avr.geo.dcmp.fb <- avr.geo[dcmp_gene, metadata_fb_dcmp_roi]
avr.geo.dcmp.fb$fb_avg <- rowMeans(avr.geo.dcmp.fb)
all(rownames(avr.geo.dcmp.cmc) == rownames(avr.geo.dcmp.endo))
avr.geo <- cbind(avr.geo.dcmp.cmc, avr.geo.dcmp.endo, avr.geo.dcmp.fb)
avr.geo <- avr.geo[,c(19,30,35)]
avr.geo.dcmp <- avr.geo

```

```{r}
Idents(dcmp) <- "Names"
dcmp <- NormalizeData(dcmp) %>% 
  FindVariableFeatures(selection.method = "vst", nfeatures = 3000) %>% 
  ScaleData()

avr.nature <- AverageExpression(dcmp, features = dcmp_gene, slot = "scale.data")
avr.nature <- avr.nature$RNA
avr.nature.sct <- rownames(avr.nature)
#avr.nature <- avr.nature[dcmp_gene,]
avr.geo.dcmp.sct <- avr.geo.dcmp[avr.nature.sct,]

ave.cor<-cor(cbind(avr.geo.dcmp.sct, avr.nature))
ave.cor <- ave.cor[1:3, 4:15]

library(corrplot)
col2 <- colorRampPalette(c("#053061","#2166AC","#4393C3","#92C5DE","#D1E5F0","#FFFFFF","#FDDBC7","#F4A582","#D6604D","#B2182B","#67001F"))
corrplot(ave.cor, col = col2(500),  method="color",order = "original", cl.lim=c(-1,1))
```

# Figure 3d
```{r}
 fgsea.dcmp.cmc <- read.csv("~/JH/HF/RData/results/fgsea_SejongIn_Clinical_phenotype_LV_FullROI_N_cellselection_Y_Cardiomyocytes_ROI_large_Vent_LV_DsSelect_DCMP__Control_Y_OL_pcaplot_outlier_RE_PID_Institute_Fixation.interval_KR_Y2023-06-22NA.csv")
  
  fgsea.icmp.cmc <- read.csv("~/JH/HF/RData/results/fgsea_SejongIn_Clinical_phenotype_LV_FullROI_N_cellselection_Y_Cardiomyocytes_ROI_large_Vent_LV_DsSelect_ICMP__Control_Y_OL_pcaplot_outlier_RE_PID_Institute_Fixation.interval_KR_Y2023-06-23NA.csv")
  
  fgsea.nes.cmc <- read.csv("~/JH/HF/RData/results/fgsea_SejongIn_Clinical_phenotype_LV_FullROI_N_cellselection_Y_Cardiomyocytes_ROI_large_Vent_LV_DsSelect_NES_HCMP__Control_Y_OL_pcaplot_outlier_RE_PID_Institute_Fixation.interval_KR_Y2023-06-22NA.csv")
  
  fgsea.es.cmc <- read.csv("~/JH/HF/RData/results/fgsea_SejongIn_Clinical_phenotype_LV_FullROI_N_cellselection_Y_Cardiomyocytes_ROI_large_Vent_LV_DsSelect_ES_HCMP__Control_Y_OL_pcaplot_outlier_RE_PID_Institute_Fixation.interval_KR_Y2023-06-22NA.csv")
  
  fgsea.nes.endo <- read.csv("~/JH/HF/RData/results/fgsea_SejongIn_Clinical_phenotype_LV_FullROI_N_cellselection_Y_Endothelial_cells_ROI_large_Vent_LV_DsSelect_NES_HCMP__Control_Y_OL_pcaplot_outlier_RE_PID_Institute_Fixation.interval_KR_Y2023-06-22NA.csv")
  
  fgsea.icmp.endo <- read.csv("~/JH/HF/RData/results/fgsea_SejongIn_Clinical_phenotype_LV_FullROI_N_cellselection_Y_Endothelial_cells_ROI_large_Vent_LV_DsSelect_ICMP__Control_Y_OL_N_RE_PID_Institute_Fixation.interval_KR_Y2023-06-22NA.csv")
  
  fgsea.dcmp.endo <- read.csv("~/JH/HF/RData/results/fgsea_SejongIn_Clinical_phenotype_LV_FullROI_N_cellselection_Y_Endothelial_cells_ROI_large_Vent_LV_DsSelect_DCMP__Control_Y_OL_pcaplot_outlier_RE_PID_Institute_Fixation.interval_KR_Y2023-06-23NA.csv")
  
  fgsea.es.endo <- read.csv("~/JH/HF/RData/results/fgsea_SejongIn_Clinical_phenotype_LV_FullROI_N_cellselection_Y_Endothelial_cells_ROI_large_Vent_LV_DsSelect_ES_HCMP__Control_Y_OL_pcaplot_outlier_RE_PID_Institute_Fixation.interval_KR_Y2023-06-22NA.csv")
  
  fgsea.list <- list()
  
  fgsea.list <- list(fgsea.dcmp.cmc, fgsea.icmp.cmc, fgsea.nes.cmc, fgsea.es.cmc, fgsea.dcmp.endo, fgsea.icmp.endo, fgsea.nes.endo, fgsea.es.endo)
  
  hallmark.list <- list()
  
  for(i in 1:length(fgsea.list)){
    hallmark.list[[i]] <- fgsea.list[[i]][fgsea.list[[i]]$padj < 0.05,] 
    hallmark.list[[i]] <- hallmark.list[[i]]$pathway
  }
  
  sig.pathway <- unlist(hallmark.list)
  sig.pathway <- unique(sig.pathway)
  
  fgsea.list.sig <- list()
  for(i in 1:length(fgsea.list)){
    fgsea.list.sig[[i]] <- fgsea.list[[i]][fgsea.list[[i]]$pathway %in% sig.pathway,] 
    fgsea.list.sig[[i]] <- fgsea.list.sig[[i]][,c(2,4,6)]
  }
  
  for(i in 1:4) {
    fgsea.list.sig[[i]]$celltype <- "Cardiomyocyte"
    fgsea.list.sig[[i+4]]$celltype <- "Endothelial cell"
  }
  
  fgsea.dcmp.cmc.sig <- fgsea.list.sig[[1]]
  fgsea.icmp.cmc.sig <- fgsea.list.sig[[2]]
  fgsea.nes.cmc.sig <- fgsea.list.sig[[3]]
  fgsea.es.cmc.sig <- fgsea.list.sig[[4]]
  fgsea.dcmp.endo.sig <- fgsea.list.sig[[5]]
  fgsea.icmp.endo.sig <- fgsea.list.sig[[6]]
  fgsea.nes.endo.sig <- fgsea.list.sig[[7]]
  fgsea.es.endo.sig <- fgsea.list.sig[[8]]
  
  fgsea.dcmp.cmc.sig$disease <- "DCMP"
  fgsea.icmp.cmc.sig$disease <- "ICMP"
  fgsea.nes.cmc.sig$disease <- "HCMPpEF"
  fgsea.es.cmc.sig$disease <- "HCMPrEF"
  fgsea.dcmp.endo.sig$disease <- "DCMP"
  fgsea.icmp.endo.sig$disease <- "ICMP"
  fgsea.nes.endo.sig$disease <- "HCMPpEF"
  fgsea.es.endo.sig$disease <- "HCMPrEF"
  
  fgsea.dcmp.cmc.sig$identity <- paste0(fgsea.dcmp.cmc.sig$disease, "_", fgsea.dcmp.cmc.sig$celltype)
  fgsea.icmp.cmc.sig$identity <- paste0(fgsea.icmp.cmc.sig$disease, "_", fgsea.icmp.cmc.sig$celltype)
  fgsea.nes.cmc.sig$identity <- paste0(fgsea.nes.cmc.sig$disease, "_", fgsea.nes.cmc.sig$celltype)
  fgsea.es.cmc.sig$identity <- paste0(fgsea.es.cmc.sig$disease, "_", fgsea.es.cmc.sig$celltype)
  fgsea.dcmp.endo.sig$identity <- paste0(fgsea.dcmp.endo.sig$disease, "_", fgsea.dcmp.endo.sig$celltype)
  fgsea.icmp.endo.sig$identity <- paste0(fgsea.icmp.endo.sig$disease, "_", fgsea.icmp.endo.sig$celltype)
  fgsea.nes.endo.sig$identity <- paste0(fgsea.nes.endo.sig$disease, "_", fgsea.nes.endo.sig$celltype)
  fgsea.es.endo.sig$identity <- paste0(fgsea.es.endo.sig$disease, "_", fgsea.es.endo.sig$celltype)
  
  fgsea.identity <- c("DCMP_Cardiomyocyte", "DCMP_Endothelial cell", "ICMP_Cardiomyocyte", "ICMP_Endothelial cell", "HCMPrEF_Cardiomyocyte", "HCMPrEF_Endothelial cell", "HCMPpEF_Cardiomyocyte", "HCMPpEF_Endothelial cell")
  
  
  fgsea.sort.sig <- rbind(fgsea.dcmp.cmc.sig, fgsea.icmp.cmc.sig, fgsea.es.cmc.sig, fgsea.nes.cmc.sig, fgsea.dcmp.endo.sig, fgsea.icmp.endo.sig, fgsea.es.endo.sig, fgsea.nes.endo.sig)
  
  fgsea.sort.sig$disease <- factor(fgsea.sort.sig$disease,
                                   levels = c("DCMP", "ICMP", "HCMPrEF", "HCMPpEF"))
  
  fgsea.sort.sig$identity <- factor(fgsea.sort.sig$identity,
                                    levels = fgsea.identity)
  
  fgsea.sort.sig$pathway <- factor(fgsea.sort.sig$pathway,
                                   levels = c("HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION", "HALLMARK_GLYCOLYSIS", "HALLMARK_MYOGENESIS", "HALLMARK_OXIDATIVE_PHOSPHORYLATION",  "HALLMARK_UV_RESPONSE_DN", "HALLMARK_APICAL_JUNCTION", "HALLMARK_TGF_BETA_SIGNALING", "HALLMARK_TNFA_SIGNALING_VIA_NFKB", "HALLMARK_ALLOGRAFT_REJECTION", "HALLMARK_INTERFERON_ALPHA_RESPONSE", "HALLMARK_INTERFERON_GAMMA_RESPONSE", "HALLMARK_KRAS_SIGNALING_UP", "HALLMARK_INFLAMMATORY_RESPONSE", "HALLMARK_IL6_JAK_STAT3_SIGNALING", "HALLMARK_IL2_STAT5_SIGNALING", "HALLMARK_ADIPOGENESIS", "HALLMARK_ESTROGEN_RESPONSE_EARLY", "HALLMARK_MYC_TARGETS_V1", "HALLMARK_FATTY_ACID_METABOLISM", "HALLMARK_COMPLEMENT", "HALLMARK_BILE_ACID_METABOLISM", "HALLMARK_ANGIOGENESIS"))
```

```{r}
  fgsea.sort.sig <- fgsea.sort.sig %>% mutate(pvalue=ifelse(padj<0.05, padj, 1))

  ggplot(fgsea.sort.sig, aes(x = identity, y = pathway, fill = NES, color = pvalue)) +
    geom_point(aes(size = -log10(padj)), alpha = 1, shape = 21, stroke = 1) +
    scale_size_continuous(range = c(0.05,7)) +
    theme_bw() +
    #theme(axis.line = element_line(colour = "black", size = 0.8)) +
    theme(panel.grid = element_blank()) +
    #ylab("Proportion of Cells") +
    xlab("") +
    #theme(text = element_text(size=11)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_y_discrete(limits = rev(levels(fgsea.sort.sig$pathway))) +
    #scale_y_continuous(expand = expansion(mult = c(0, 0))) +
    #guides(fill = guide_legend(ncol = 1)) +
    #theme(legend.key.height = unit(10, units = 'pt')) +
    scale_fill_gradientn(colors = brewer.pal(n = 11, name = "RdBu")[11:1], limits = c(min(fgsea.sort.sig$NES), max(fgsea.sort.sig$NES))) + 
    scale_color_manual(values = c("black", "white")) + 
    geom_vline(xintercept = xintercepts, 
               color = "black", linetype = "dashed")
  
  
  xintercepts <- seq(2.5, length(unique(fgsea.sort.sig$identity)), by = 2)
```

# Figure 3e
```{r}
cmc.es.nes <- read.csv("~/JH/HF/RData/results/SejongIn_Clinical_phenotype_LV_FullROI_N_cellselection_Y_Cardiomyocytes_ROI_large_Vent_LV_DsSelect_ES_HCMP__NES_HCMP_Y_OL_pcaplot_outlier_RE_PID_Fixation.interval_KR_Y2023-06-23conditionES_HCMP-conditionNES_HCMP.csv")
ec.es.nes <- read.csv("~/JH/HF/RData/results/SejongIn_Clinical_phenotype_LV_FullROI_N_cellselection_Y_Endothelial_cells_ROI_large_Vent_LV_DsSelect_ES_HCMP__NES_HCMP_Y_OL_pcaplot_outlier_RE_PID_Fixation.interval_KR_Y2023-06-23conditionES_HCMP-conditionNES_HCMP.csv")

## volcano ----
de <- ec.es.nes
de$logFC <- as.numeric(de$logFC)
de$adj.P.Val <- as.numeric(de$adj.P.Val)
de$diffexpressed <- "NO"    # UP DOWN GENE을 유의미한 범위서 구분하기 위한 방법 
de$diffexpressed[de$logFC >  0.5 & de$adj.P.Val < 0.05] <- "UP"           
de$diffexpressed[de$logFC < -0.5 & de$adj.P.Val < 0.05] <- "DOWN" #기분 나빠 보이는 gene은 다 날려버리기
de$delabel <- NA
de$delabel[de$diffexpressed != "NO"] <- de$SYMBOL[de$diffexpressed != "NO"]     # 유의미한 gene만 labeling 하기 

de$adj.P.Val[de$adj.P.Val < 5e-320 ] <- 5e-320     # p_val이 0인 애들이 border에서 구리구리하게 나오는걸 방지하기 위해 임의의 값을 부여 (나중에 0임을 알려줘야 하긴 함)

#"Control" = "#A8A196", "DCMP" = "#E14D2A", "ICMP" = "#FD841F", "HCMPrEF" = "#10A19D", "HCMPpEF" = "#3E6D9C"

ggplot(data=de, aes(x=logFC, y=-log10(adj.P.Val), col=diffexpressed, label=delabel)) +
  geom_point() + 
  theme_classic()+
  geom_text_repel(size=5, show.legend = F,min.segment.length = 1, nudge_x = c(0.1, -0.1) ,
                  nudge_y = c(5), color= "black", fontface = "italic", fontfamily="arial")+
  scale_color_manual(values=c("#3E6D9C", "grey", "#10A19D")) +
  geom_vline(xintercept=c(-0.25, 0.25), col="grey", linetype='dashed') +
  geom_hline(yintercept=-log10(0.05), col="grey", linetype='dashed') +
  xlim(-max(abs(de$logFC)), max(abs(de$logFC))) 

  ggplot(data=de, aes(x=logFC, y=-log10(adj.P.Val), col=diffexpressed)) +
  geom_point() + 
  theme_classic()+
  scale_color_manual(values=c("#3E6D9C", "grey", "#10A19D")) +
  geom_vline(xintercept=c(-0.25, 0.25), col="grey", linetype='dashed') +
  geom_hline(yintercept=-log10(0.05), col="grey", linetype='dashed') +
  xlim(-max(abs(de$logFC)), max(abs(de$logFC))) + NoLegend()
```

# Figure 3g
```{r}
tmp_df_tidy_pval<- read.csv("~/JH/HF/RData/results/fgsea_SejongIn_Clinical_phenotype_LV_FullROI_N_cellselection_Y_Cardiomyocytes_ROI_large_Vent_LV_DsSelect_ES_HCMP__NES_HCMP_Y_OL_pcaplot_outlier_RE_PID_Fixation.interval_KR_Y2023-06-23NA.csv")

  picture_pathway <- ggplot(tmp_df_tidy_pval, aes(reorder(pathway, NES), NES)) +
    geom_col(aes(fill=factor(clas))) +
    scale_fill_manual(values = c("gray67", "#3E6D9C", "#10A19D")) +
    
    coord_flip() +
    labs(x="", y="Normalized Enrichment Score",
         title=paste(gsub("ES_HCMP", "HCMPrEF", gsub("NES_HCMP", "HCMPpEF", gsub("condition", "", i))), "in", a)) + 
    theme_minimal() +
    theme(legend.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(color = "black", fill = NA),
          axis.line.x = element_line(color = "black"),
          axis.ticks.x = element_line(color = "black"),
          axis.ticks.length.x = unit(0.3, "cm"))
```

